---
title: "Homework 4"
output: 
  html_document:
    keep_md: yes
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(gapminder)
library(dplyr)
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

# Tasks :
##### 1. Putting data in wider format;
##### 2. Producing a plot of the wide data; (10%)
##### 3. Re-lengthening the wider data. (10%)
 

## Exercise 1: 
#### Univariate Option 1

#### Make a tibble with one row per year, and columns for life expectancy for two or more countries.
```{r}
(exer1 = gapminder %>% 
  filter(country == "Canada" | country == "China" | country == "Angola") %>% 
     pivot_wider(id_cols = c(year), 
              names_from = country, 
              values_from = lifeExp)) %>% 
  DT :: datatable(exer1)
 
```

#### Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
exer1 %>% 
  ggplot(aes(year, Angola, colour = "Angola")) +
    geom_point( ) +
    geom_point(aes(year, Canada, colour = "Canada"))+
    geom_point(aes(year, China, colour = "China")) +
    xlab('Year') +
    ylab('Life Expectancy')
```

#### Re-lengthen the data.

```{r}
exer1 %>%  
  pivot_longer(cols = Angola:China, 
               names_to  = "Country", 
               values_to = "Life Expectancy") %>% 
DT:: datatable(exer1)
```

#Exercise 2: Multivariate Data Reshaping

### Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries

#### Putting data in wider format;
```{r}
(exer2 = gapminder %>% 
  filter(country == 'Canada' | country == 'China' | country == 'Senegal') %>% 
  pivot_wider(id_cols = c(year),
              names_from = c(country),
              names_sep = "_",
              values_from = c(lifeExp,gdpPercap))) %>% 
  DT :: datatable(exer2)
              
```

#### Re-lengthening the data

```{r}
exer2 %>% 
 pivot_longer(cols      = -year, 
               names_to  = c(".value","Country"),
               names_sep = "_") %>% 
  DT::datatable(exer2)
```

# Exercise 3: Table Joins 

### Copy and Paste the following: 
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")

```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```


## Exercise 3.1: For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

```{r}
(new_email = email %>%
   separate_rows(guest, sep = ", ") %>% 
   rename(name = guest))
guest %>% 
  inner_join(new_email, by = 'name')
```

## Exercise 3.2: Who do we have emails for, yet are not on the guestlist?

```{r}
new_email %>% 
  anti_join(guest, by = 'name')
```


## Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

```{r}
(new_guestlist = new_email %>% 
  full_join(guest, by = "name") %>% 
  filter(email != 'NA')) %>% 
  select(name, email)
```

